Multilevel regression models

### 2025-09-30 processing/prod_analysis-ocde-sample.Rmd
to-ignore2

>>>>

# Numeric -  Egalitarian Attitudes Index

Cronbach's alpha is 0.7 for:

1. It is the responsibility of the government to reduce the differences in income between people with high incomes and those with low incomes. 
2. For a society to be fair, differences in people's standard of living should be small.

Note: the item "Differences in income in [COUNTRY] are too large" is not included because of two reasons. First, the item is fuzzy in terms of what is it measuring, "are" refers to a perception about inequality and "too large" as a normative judgement. Second, the literature has addressed this indicator as perceived or concerns about inequality.

<!----------------------------------------------------------------------------->

```{r mod-homclass-numeric,results='asis'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,sjlabelled,haven,stringr,ordinal)  
rm(list=ls())
load(here::here("input/data/proc/study1_country.RData"));df1 <- df2
dfreg <- df1 %>% dplyr::select(
                        egal=egal3,
                        class3,
                        homclass,
                        Q03pcm,
                        edyears,
                        female,
                        agenum,
                        union,
                        workst,
                        WEIGHT,
                        region,
                        "gini_disp",
                        "gini_mkt",
                        rgdpna,
                        oecd,
                        country2,country
                        ) %>% 
  mutate(logrgdpna=log(rgdpna),
         edyears2=edyears^2,
         age2=agenum^2) %>%
 filter(oecd=="OECD" & country2 != "SVN") %>% 
  na.omit()

# Models 
base <- lmer(egal~1 + (1|country2),data=dfreg,weights = WEIGHT); icc1<- performance::icc(base)
homclass<- update(base, . ~ . +homclass)
full1 <- update(homclass, . ~ . + class3+female+agenum+age2)
rob1 <- update(full1, . ~ . + class3+edyears+ Q03pcm+union+workst)
rob2 <- update(rob1, . ~ . -(1|country2) + (homclass|country2))
# anova(rob1,rob2)
# Interactions segregation x class
int_homo <- update(rob1, . ~ . +class3*homclass)

models <- list(homclass,full1,rob1,int_homo)
save(models,file = here::here("output/ocde-egalitarian-homclass-full-individual.RData"),compress = "gzip")
```

```{r mod-intermacro-numeric,results='asis'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,sjlabelled,haven,stringr,ordinal)  
rm(list=ls())
load(here::here("input/data/proc/study1_country.RData"));df1 <- df2
dfreg <- df1 %>% dplyr::select(egal=egal2,
                               class3,
                        homclass,
                        Q03pcm,
                        edyears,
                        female,
                        agenum,
                        union,
                        workst,
                        WEIGHT,
                        region,
                        "gini_disp",
                        "gini_mkt",
                        rgdpna,
                        oecd,
                        country2,country,
                        palmaratio,
                        top10,
                        middle50,
                        d10d1,
                        giniindex,
                        ttheilge1
                        ) %>% 
  mutate(logrgdpna=log(rgdpna),
         edyears2=edyears^2,
         age2=agenum^2) %>%
  filter(oecd=="OECD" & country2 != "SVN") %>% 
  na.omit()

# Homoclass x Gini--------------------------------------------------------------
main <- "egal~class3+homclass+edyears+Q03pcm+workst+union+female+agenum+age2+(1|country2)"
base <- lmer(formula(main),data=dfreg,weights = WEIGHT)
giniD_gdp<- update(base, . ~ . +gini_disp+logrgdpna)
giniD_rs <- update(giniD_gdp, . ~ . - (1|country2) + (homclass|country2))
# anova(giniD_gdp,giniD_rs)
giniD_int<- update(giniD_rs, . ~ . + homclass*gini_disp + (homclass|country2))
# summary(giniD_int)

giniM_gdp<- update(base, . ~ . +gini_mkt+logrgdpna)
giniM_rs <- update(giniM_gdp, . ~ . - (1|country2) + (homclass|country2))
# anova(giniM_gdp,giniM_rs)
giniM_int <- update(giniM_rs, . ~ . + homclass*gini_mkt + (homclass|country2))
# summary(giniM_int)
models <- list(giniD_gdp,giniD_rs,giniD_int,
               giniM_gdp,giniM_rs,giniM_int)

# Homoclass x Other Economic Inequality-----------------------------------------
main <- "egal~class3+homclass+edyears+Q03pcm+workst+union+female+agenum+age2+logrgdpna+(homclass|country2)"
base <- lmer(formula(main),data=dfreg,weights = WEIGHT)
int_palma <- update(base, . ~ . +homclass*palmaratio)
int_pc10  <- update(base, . ~ . +homclass*top10)
int_pc50  <- update(base, . ~ . +homclass*middle50)
int_d10d1 <- update(base, . ~ . +homclass*d10d1)
int_gini_wiid <- update(base, . ~ . +homclass*giniindex)
int_theil  <- update(base, . ~ . +homclass*ttheilge1)

models_m <- list(int_palma,int_pc10,int_pc50,int_d10d1,
                 int_gini_wiid,int_theil
                 )
save(giniD_gdp,giniD_rs,giniD_int,
                 giniM_gdp,giniM_rs,giniM_int,
                 int_palma,int_pc10,int_pc50,int_d10d1,
                 int_gini_wiid,int_theil,
     file = here::here("output/ocde-egalitarian-homclass-full-macro.RData"),compress = "gzip")
```

```{r}
load(here::here("output/ocde-egalitarian-homclass-full-individual.RData"))
texreg::knitreg(models)

load(here::here("output/ocde-egalitarian-homclass-full-macro.RData"))
texreg::knitreg(list(giniD_int,giniM_int,int_palma,int_pc10,int_pc50,int_d10d1,
                 int_gini_wiid,int_theil))
```

```{r mod-homclass-numeric-strong-ties,results='asis'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,sjlabelled,haven,stringr,ordinal)  
rm(list=ls())
load(here::here("input/data/proc/study1_country.RData"));df1 <- df2
dfreg <- df1 %>% dplyr::select(egal=egal2,
                        class3,
                        homclass=homclass_s,
                        Q03pcm,
                        edyears,
                        female,
                        agenum,
                        union,
                        workst,
                        WEIGHT,
                        region,
                        "gini_disp",
                        "gini_mkt",
                        rgdpna,
                        oecd,
                        country2,country
                        ) %>% 
  mutate(logrgdpna=log(rgdpna),
         edyears2=edyears^2,
         age2=agenum^2) %>%
  filter(oecd=="OECD" & country2 != "SVN") %>%   
  na.omit()

# Models 
base <- lmer(egal~1 + (1|country2),data=dfreg,weights = WEIGHT); icc1<- performance::icc(base)
homclass<- update(base, . ~ . +homclass)
full1 <- update(homclass, . ~ . + class3+female+agenum+age2)
rob1 <- update(full1, . ~ . + class3+edyears+ Q03pcm+union+workst)
rob2 <- update(rob1, . ~ . -(1|country2) + (homclass|country2))
# anova(rob1,rob2)
# Interactions segregation x class
int_homo <- update(rob1, . ~ . +class3*homclass)

models <- list(homclass,full1,rob1,int_homo)
save(models,file = here::here("output/ocde-egalitarian-homclass-strong-individual.RData"),compress = "gzip")
```

```{r mod-intermacro-numeric-strong-ties,results='asis'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,sjPlot,sjlabelled,haven,stringr,gridExtra)  
rm(list=ls())
load(here::here("input/data/proc/study1_country.RData"));df1 <- df2
dfreg <- df1 %>% dplyr::select(egal=egal2,
                        class3,
                        homclass=homclass_s,
                        Q03pcm,
                        edyears,
                        female,
                        agenum,
                        union,
                        workst,
                        WEIGHT,
                        region,
                        "gini_disp",
                        "gini_mkt",
                        rgdpna,
                        oecd,
                        country2,country,
                        palmaratio,
                        top10,
                        middle50,
                        d10d1,
                        giniindex,
                        ttheilge1
                        ) %>% 
  mutate(logrgdpna=log(rgdpna),
         edyears2=edyears^2,
         age2=agenum^2) %>%
filter(oecd=="OECD" & country2 != "SVN") %>% 
  na.omit()

# Homoclass x Gini--------------------------------------------------------------
main <- "egal~class3+homclass+edyears+Q03pcm+workst+union+female+agenum+age2+(1|country2)"
base <- lmer(formula(main),data=dfreg,weights = WEIGHT)
giniD_gdp<- update(base, . ~ . +gini_disp+logrgdpna)
giniD_rs <- update(giniD_gdp, . ~ . - (1|country2) + (homclass|country2))
# anova(giniD_gdp,giniD_rs)
giniD_int<- update(giniD_rs, . ~ . + homclass*gini_disp + (homclass|country2))
summary(giniD_int)

# Homoclass x Other Economic Inequality-----------------------------------------
main <- "egal~class3+homclass+edyears+Q03pcm+workst+union+female+agenum+age2+logrgdpna+(homclass|country2)"
base <- lmer(formula(main),data=dfreg,weights = WEIGHT)
int_palma <- update(base, . ~ . +homclass*palmaratio)
int_pc10  <- update(base, . ~ . +homclass*top10)
int_pc50  <- update(base, . ~ . +homclass*middle50)
int_d10d1 <- update(base, . ~ . +homclass*d10d1)
int_gini_wiid <- update(base, . ~ . +homclass*giniindex)
int_theil  <- update(base, . ~ . +homclass*ttheilge1)

models_m <- list(int_palma,int_pc10,int_pc50,int_d10d1,
                 int_gini_wiid,int_theil
                 )
save(giniD_gdp,giniD_rs,giniD_int,
                 giniM_gdp,giniM_rs,giniM_int,
                 int_palma,int_pc10,int_pc50,int_d10d1,
                 int_gini_wiid,int_theil,
     file = here::here("output/ocde-egalitarian-homclass-strong-macro.RData"),compress = "gzip")
```

```{r}
load(here::here("output/ocde-egalitarian-homclass-strong-individual.RData"))
texreg::knitreg(models)

load(here::here("output/ocde-egalitarian-homclass-strong-macro.RData"))
texreg::knitreg(list(giniD_int,giniM_int,int_palma,int_pc10,int_pc50,int_d10d1,
                 int_gini_wiid,int_theil))
```

```{r mod-homclass-numeric-strong-ties,results='asis'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,sjlabelled,haven,stringr,ordinal)  
rm(list=ls())
load(here::here("input/data/proc/study1_country.RData"));df1 <- df2
dfreg <- df1 %>% dplyr::select(egal=egal2,
                        class3,
                        homclass=homclass_w,
                        Q03pcm,
                        edyears,
                        female,
                        agenum,
                        union,
                        workst,
                        WEIGHT,
                        region,
                        "gini_disp",
                        "gini_mkt",
                        rgdpna,
                        oecd,
                        country2,country
                        ) %>% 
  mutate(logrgdpna=log(rgdpna),
         edyears2=edyears^2,
         age2=agenum^2) %>%
filter(oecd=="OECD" & country2 != "SVN") %>% 
  na.omit()

# Models 
base <- lmer(egal~1 + (1|country2),data=dfreg,weights = WEIGHT); icc1<- performance::icc(base)
homclass<- update(base, . ~ . +homclass)
full1 <- update(homclass, . ~ . + class3+female+agenum+age2)
rob1 <- update(full1, . ~ . + class3+edyears+ Q03pcm+union+workst)
rob2 <- update(rob1, . ~ . -(1|country2) + (homclass|country2))
# anova(rob1,rob2)
# Interactions segregation x class
int_homo <- update(rob1, . ~ . +class3*homclass)

models <- list(homclass,full1,rob1,int_homo)
save(models,file = here::here("output/ocde-egalitarian-homclass-weak-individual.RData"),compress = "gzip")
```

```{r mod-intermacro-numeric-strong-ties,results='asis'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,sjPlot,sjlabelled,haven,stringr,gridExtra)  
rm(list=ls())
load(here::here("input/data/proc/study1_country.RData"));df1 <- df2
dfreg <- df1 %>% dplyr::select(egal=egal2,
                        class3,
                        homclass=homclass_w,
                        Q03pcm,
                        edyears,
                        female,
                        agenum,
                        union,
                        workst,
                        WEIGHT,
                        region,
                        "gini_disp",
                        "gini_mkt",
                        rgdpna,
                        oecd,
                        country2,country,
                        palmaratio,
                        top10,
                        middle50,
                        d10d1,
                        giniindex,
                        ttheilge1
                        ) %>% 
  mutate(logrgdpna=log(rgdpna),
         edyears2=edyears^2,
         age2=agenum^2) %>%
  filter(oecd=="OECD" & country2 != "SVN") %>% 
  na.omit()

# Homoclass x Gini--------------------------------------------------------------
main <- "egal~class3+homclass+edyears+Q03pcm+workst+union+female+agenum+age2+(1|country2)"
base <- lmer(formula(main),data=dfreg,weights = WEIGHT)
giniD_gdp<- update(base, . ~ . +gini_disp+logrgdpna)
giniD_rs <- update(giniD_gdp, . ~ . - (1|country2) + (homclass|country2))
# anova(giniD_gdp,giniD_rs)
giniD_int<- update(giniD_rs, . ~ . + homclass*gini_disp + (homclass|country2))
summary(giniD_int)

# Homoclass x Other Economic Inequality-----------------------------------------
main <- "egal~class3+homclass+edyears+Q03pcm+workst+union+female+agenum+age2+logrgdpna+(homclass|country2)"
base <- lmer(formula(main),data=dfreg,weights = WEIGHT)
int_palma <- update(base, . ~ . +homclass*palmaratio)
int_pc10  <- update(base, . ~ . +homclass*top10)
int_pc50  <- update(base, . ~ . +homclass*middle50)
int_d10d1 <- update(base, . ~ . +homclass*d10d1)
int_gini_wiid <- update(base, . ~ . +homclass*giniindex)
int_theil  <- update(base, . ~ . +homclass*ttheilge1)

models_m <- list(int_palma,int_pc10,int_pc50,int_d10d1,
                 int_gini_wiid,int_theil
                 )
save(giniD_gdp,giniD_rs,giniD_int,
                 giniM_gdp,giniM_rs,giniM_int,
                 int_palma,int_pc10,int_pc50,int_d10d1,
                 int_gini_wiid,int_theil,
     file = here::here("output/ocde-egalitarian-homclass-weak-macro.RData"),compress = "gzip")
```

```{r}
load(here::here("output/ocde-egalitarian-homclass-weak-individual.RData"))
texreg::knitreg(models)

load(here::here("output/ocde-egalitarian-homclass-weak-macro.RData"))
texreg::knitreg(list(giniD_int,giniM_int,int_palma,int_pc10,int_pc50,int_d10d1,
                 int_gini_wiid,int_theil))
```
## R's class

```{r mod-homclass-numeric-respondent,results='asis'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,sjlabelled,haven,stringr,ordinal)  
rm(list=ls())
load(here::here("input/data/proc/study1_country.RData"));df1 <- df2
dfreg <- df1 %>% dplyr::select(egal=egal2,
                        class3=class3res,
                        homclass=homclass_res,
                        class3spo,
                        Q03pcm,
                        edyears,
                        female,
                        agenum,
                        union,
                        workst,
                        WEIGHT,
                        region,
                        "gini_disp",
                        "gini_mkt",
                        rgdpna,
                        oecd,
                        country2,country
                        ) %>% 
  mutate(logrgdpna=log(rgdpna),
         edyears2=edyears^2,
         age2=agenum^2) %>%
  filter(oecd=="OECD" & country2 != "SVN")

# Models 
base <- lmer(egal~1 + (1|country2),data=dfreg,weights = WEIGHT); icc1<- performance::icc(base)
homclass<- update(base, . ~ . +homclass)
full1 <- update(homclass, . ~ . + class3+female+agenum+age2)
rob1 <- update(full1, . ~ . + class3+edyears+ Q03pcm+union+workst+class3spo)
rob2 <- update(rob1, . ~ . -(1|country2) + (homclass|country2))
# anova(rob1,rob2)
# Interactions segregation x class
int_homo <- update(rob1, . ~ . +class3*homclass)

models <- list(homclass,full1,rob1,int_homo)
save(models,file = here::here("output/ocde-egalitarian-homclass-respondent.RData"),compress = "gzip")
```

```{r mod-intermacro-numeric-respondent,results='asis'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,sjPlot,sjlabelled,haven,stringr,gridExtra)  
rm(list=ls())
load(here::here("input/data/proc/study1_country.RData"));df1 <- df2
dfreg <- df1 %>% dplyr::select(egal=egal2,
                        class3=class3res,
                        class3spo,
                        homclass=homclass_res,
                        Q03pcm,
                        edyears,
                        female,
                        agenum,
                        union,
                        workst,
                        WEIGHT,
                        region,
                        "gini_disp",
                        "gini_mkt",
                        rgdpna,
                        oecd,
                        country2,country,
                        palmaratio,
                        top10,
                        middle50,
                        d10d1,
                        giniindex,
                        ttheilge1
                        ) %>% 
  mutate(logrgdpna=log(rgdpna),
         edyears2=edyears^2,
         age2=agenum^2) %>%
  filter(oecd=="OECD" & country2 != "SVN")

# Homoclass x Gini--------------------------------------------------------------
main <- "egal~class3+homclass+edyears+Q03pcm+workst+union+female+agenum+age2+class3spo+(1|country2)"
base <- lmer(formula(main),data=dfreg,weights = WEIGHT)
giniD_gdp<- update(base, . ~ . +gini_disp+logrgdpna)
giniD_rs <- update(giniD_gdp, . ~ . - (1|country2) + (homclass|country2))
# anova(giniD_gdp,giniD_rs)
giniD_int<- update(giniD_rs, . ~ . + homclass*gini_disp + (homclass|country2))
summary(giniD_int)

# Homoclass x Other Economic Inequality-----------------------------------------
main <- "egal~class3+homclass+edyears+Q03pcm+workst+union+female+agenum+age2+logrgdpna+ class3spo +(homclass|country2)"
base <- lmer(formula(main),data=dfreg,weights = WEIGHT)
int_palma <- update(base, . ~ . +homclass*palmaratio)
int_pc10  <- update(base, . ~ . +homclass*top10)
int_pc50  <- update(base, . ~ . +homclass*middle50)
int_d10d1 <- update(base, . ~ . +homclass*d10d1)
int_gini_wiid <- update(base, . ~ . +homclass*giniindex)
int_theil  <- update(base, . ~ . +homclass*ttheilge1)

models_m <- list(int_palma,int_pc10,int_pc50,int_d10d1,
                 int_gini_wiid,int_theil
                 )
save(giniD_gdp,giniD_rs,giniD_int,
                 giniM_gdp,giniM_rs,giniM_int,
                 int_palma,int_pc10,int_pc50,int_d10d1,
                 int_gini_wiid,int_theil,
     file = here::here("output/ocde-egalitarian-homclass-respondent-macro.RData"),compress = "gzip")
```

```{r}
load(here::here("output/ocde-egalitarian-homclass-respondent.RData"))
texreg::knitreg(models)

load(here::here("output/ocde-egalitarian-homclass-respondent-macro.RData"))
texreg::knitreg(list(giniD_int,giniM_int,int_palma,int_pc10,int_pc50,int_d10d1,
                 int_gini_wiid,int_theil))

marginaleffects::plot_predictions(giniD_int, condition = list("homclass",gini_disp=range),gray = F)
marginaleffects::plot_predictions(int_gini_wiid, condition = list("homclass",giniindex=range),gray = F)
marginaleffects::plot_predictions(giniM_int, condition = list("homclass",gini_mkt=range),gray = F)
```
